Data Specialization for a Staged Lambda Calculus

### 2015-02-05 paper/splitting.tex
few small edits

In this section, we provide an algorithm to translate the multistage terms and values of \lang\
into equivalent but simpler forms where the stage \bbone\ and stage \bbtwo\ portions 
of the computation are {\em syntactically seperated} from each other.
Specifically, we desire a outputs of the form $\mathcal{C}(e_1;e_2)$ where  
$e_1$ is a monostage term that encodes all stage \bbone\ computation, 
and $e_2$ is a monostage term that encodes all stage \bbtwo\ computation.
To represent outputs in this way, we introduce three new constructs:
\begin{itemize}
\item {\em Single-output pipelines}, \texttt{pipeS}, for the output of splitting \lang\ terms at world \bbtwo.
\item {\em Multi-output pipelines}, \texttt{pipeM}, for the output of splitting \lang\ terms at world \bbonem. 
\item {\em Masked values}, \texttt{mv}, which will serve a purpose analogous to \lang\ partial values.
\end{itemize}

We also define analogs to evaluation and residualization for these new terms, 
called $\sepredonesym$ and $\sepredtwosym$.
So, whereas residualization ($\redtwosym$) sends multistage terms at world \bbtwo\ to residuals, 
separated-form residualization ($\sepredtwosym$) sends \texttt{pipeS} terms to residuals.
And whereas evaluation ($\redonesym$) sends multistage terms at world \bbonem\ to partial values,
seperated-form evaluation ($\sepredonesym$) sends a \texttt{pipeM} terms to \texttt{mv} terms.
An overview of these features is provided in \ref{fig:outputTermSummary}.
That figure also provides the actual rules for $\sepredonesym$ and $\sepredtwosym$.
Note how there's one rule for each, and they're rather simple.
Indeed, the simplicity of these reduction rules is one reason why we 
would want to translate into this separated form, rather than use the semantics of \ref{sec:semantics}.

@@

Exactly what is meant by ``such that the semantics is preserved'' depends on whether we're translating
terms at \bbonem\ or \bbtwo.  
In the latter case, where an $\coltwo e A$ is translated to a $\pipeS p l r$ with the $\splitonesym$ relation,
this constraint is straightforward: $e$ and $\pipeS p l r$ can both be reduced to residuals
via $\redtwosym$ and $\sepredtwosym$, so we only need to require that those
residuals be equivalent for the translation to be correct.
In the former case, where an $\colone e A$ is translated to a $\pipeM c l r$ with the $\splittwosym$ relation, 
things are more complicated: 
the output of $\redonesym$ is a partial value and the output of $\sepredonesym$ is a masked value,
so we first need a way to relate those forms if we want to talk about correctness of the term translation.
That is the purpose of of the {\em masking} relation, written $\vsplito$.

>>>>

In this section, we provide an algorithm to translate the multistage terms and values of \lang\
into equivalent but simpler forms where the stage \bbone\ and stage \bbtwo\ portions 
of the computation are {\em syntactically seperated} from each other.
That is, we desire a outputs of the form $\mathcal{C}(e_1;e_2)$ where  
$e_1$ is a monostage term that encodes all stage \bbone\ computation, 
and $e_2$ is a monostage term that encodes all stage \bbtwo\ computation.
To represent outputs in this way, we introduce three new constructs:
\begin{itemize}
\item {\em Single-output pipelines}, \texttt{pipeS}, for the output of splitting \lang\ terms at world \bbtwo.
\item {\em Multi-output pipelines}, \texttt{pipeM}, for the output of splitting \lang\ terms at world \bbonem. 
\item {\em Masked values}, \texttt{mv}, which will serve a purpose analogous to \lang\ partial values.
\end{itemize}

We also define analogs to evaluation and residualization for these new terms, 
called $\sepredonesym$ and $\sepredtwosym$, 
which respectively send \texttt{pipeM} terms to \texttt{mv} terms
and \texttt{pipeS} terms to residuals.
An overview of these features is provided in \ref{fig:outputTermSummary}.
That figure also provides the actual rules for $\sepredonesym$ and $\sepredtwosym$.
Note how there's one rule for each, and they're rather simple.
Indeed, the simplicity of these reduction rules is one reason why we 
would want to translate into this separated form, rather than use the semantics of \ref{sec:semantics}.

@@

Exactly what is meant by ``such that the semantics is preserved'' depends on whether we're translating
terms at \bbonem\ or \bbtwo.  
In the latter case, where an $\coltwo e A$ is translated to a $\pipeS p l r$ with the $\splittwosym$ relation,
this constraint is straightforward: $e$ and $\pipeS p l r$ can both be reduced to residuals
via $\redtwosym$ and $\sepredtwosym$, so we only need to require that those
residuals be equivalent for the translation to be correct.
In the former case, where an $\colone e A$ is translated to a $\pipeM c l r$ with the $\splitonesym$ relation, 
things are more complicated: 
the output of $\redonesym$ is a partial value and the output of $\sepredonesym$ is a masked value,
so we first need a way to relate those forms if we want to talk about correctness of the term translation.
That is the purpose of of the {\em masking} relation, written $\vsplito$.
